---
title: "Project3: Redo of Project 2 Predicting Species of Microtus"
subtitle: "STAT 601: Group 5 (Fall 2024)"
author: "Neha Karna, Prafulla Shrestha, 
Aidan Stewart, Josh Lefdal, Shivam Bhardwaj"
date: "3rd Dec, 2024"
output:
  pdf_document: default
  html_document: default
  keep_md: TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning=FALSE, message=FALSE, fig.width = 10, fig.height = 7, dev = "jpeg",
  fig.path = "plots/")
```

```{r, warning=FALSE, message=FALSE}

library(dplyr)
library(GGally)

```

# Data Manipulation for known dataset

1.  Read the Data "Vole Skulls.xlsm"
2.  Create three datasets: "Unknown", "Subterraneus", "Multiplex"
3.  Change column names for consistency through all three datasets

```{r}
library(readxl)
#Read data files into R
file_path <- "Vole Skulls.xlsm"

unknown <- read_excel(file_path, sheet = "Unknown")
subterraneus <- read_excel(file_path, sheet = "Subterraneus")
multiplex <- read_excel(file_path, sheet = "Multiplex")

# change column names for consistency
names(unknown) <- c("index", "chromosomal_id", "skull_length", "skull_height", "skull_width")
names(subterraneus) <- names(unknown)
names(multiplex) <- names(unknown)

```

4.  Remove NA rows in Subterraneus and Multiplex datasets
5.  Merge "Subterraneus" and "Multiplex" and name it "Known"
6.  Create response variable.
7.  Add response variable where 1 represents subterraneus and 0 represents Multiplex

```{r}
# Remove NA values for subterraneus and multiplex data sets
subterraneus <- subterraneus[!apply(subterraneus, 1, function(x) {any(is.na(x))}), ]
multiplex <- multiplex[!apply(multiplex, 1, function(x) {any(is.na(x))}), ]

# Combine the subterraneus and multiplex data sets
known<- rbind(subterraneus, multiplex)

# Reset the "Index" to count from 1 to the total number of rows
known <- known %>%
  mutate(`index` = row_number())

# Assign a binary response variable to the Known data
known$response <- ifelse(known$`chromosomal_id` == "subterraneus", 1, 0)
```

### Visualize known data set for chromosomal id

Observe the known data set outliers in a table.

```{r}
# Load the tidyr package if not already loaded
library(tidyr)

# Reshape using pivot_longer
known_long <- known %>%
  pivot_longer(
    cols = c(skull_length, skull_height, skull_width),  # Columns to reshape
    names_to = "measurement",                          # New column for measurement type
    values_to = "value"                                # New column for measurement values
  )
# Identify outliers based on IQR
outliers <- known_long %>%
  group_by(measurement) %>%
  filter(value < quantile(value, 0.25) - 1.5 * IQR(value) |
         value > quantile(value, 0.75) + 1.5 * IQR(value))

print(paste("List of Outliers in the Known Dataset"))
print(outliers)

library(ggplot2)
#Boxplot comparison in a single plot with labled outliers 
ggplot(known_long, aes(x = chromosomal_id, y = value, fill=chromosomal_id)) +
  geom_boxplot() +
  geom_text(data = outliers, aes(label = index), color = "red", size = 2.5, vjust = -0.5) + #lable outlier with index value
  facet_wrap(~ measurement, scales = "free",
             labeller = labeller(measurement = c("skull_length" = "Vole Skull Length",
                                                 "skull_height" = "Vole Skull height",
                                                 "skull_width" = "Vole Skull width"))) +
  labs(title = "Boxplot of Vole Skull Measurements by Chromosomal Id",
       x = "Chromosomal Id",
       y = "Measurement (0.01mm)") +
  theme_minimal()

#Histogram comparison in a single plot
ggplot(known_long, aes(x = value, fill = chromosomal_id)) +
  geom_histogram(position = "identity", alpha = 0.6, bins = 30) +  
  facet_wrap(~ measurement, scales = "free", 
             labeller = labeller(measurement = c("skull_length" = "Vole Skull Length",
                                                 "skull_height" = "Vole Skull Height",
                                                 "skull_width" = "Vole Skull Width"))) +
  labs(title = "Histogram of Vole Skull Measurements by Chromosomal Id",
       x = "Measurement (0.01mm)",
       y = "Frequency") +
  theme_minimal() +
  scale_fill_manual(values = c("subterraneus" = "blue", "multiplex" = "red"))
```

## Remove outliers

Removing the rows with outliers in any of the columns

```{r}
# Identify outliers based on IQR
outliers <- known_long %>%
  group_by(measurement) %>%
  filter(value < quantile(value, 0.25) - 1.5 * IQR(value) |
         value > quantile(value, 0.75) + 1.5 * IQR(value))

# Check and filter outliers based on these thresholds
outliers <- known_long %>%
  group_by(measurement) %>%
  filter(value < quantile(value, 0.25) - 1.5 * IQR(value) |
         value > quantile(value, 0.75) + 1.5 * IQR(value))

# Remove rows in the 'known' dataset with any outlier
known <- known %>%
  filter(!index %in% outliers$index)

# Rename 'value' column to 'outlier_value'
outliers <- outliers %>%
  rename(outlier_value = value) %>%
  select(-response)

cat("\nIdentified Outliers:\n")
print(outliers)
```

# EDA for known dataset

## Descriptive Summary

Obtaining a descripitive summary of the known dataset following removal of outliers

```{r}

#to obtain Descriptive Summary with clean format
summary_df<- known %>%
  group_by(chromosomal_id) %>%
  summarize(
    N= n(),
    min_skull_length = min(skull_length),
    max_skull_length = max(skull_length),
    mean_skull_length = mean(skull_length),
    min_skull_height = min(skull_height),
    max_skull_height = max(skull_height),
    mean_skull_height = mean(skull_height),
    min_skull_width = min(skull_width),
    max_skull_width= max(skull_width),
    mean_skull_width = mean(skull_width)
  )
#rename for better readability
colnames(summary_df) <- c("Chromosomal type","number", 
                          "minimum skull length", "maximum skull length", "mean skull length",
                          "minimum skull height", "maximum skull height", "mean skull height",
                          "minimum skull width", "maximum skull width", "mean skull width"
                          )
#transpose to view the table in long format
summary_df<- as.data.frame(t(summary_df)) 
colnames(summary_df) <- NULL

cat("\nDescriptive Summary:\n")
print(summary_df)

```

## Box plot

Observe the known dataset in a boxplot, confirming outliers were removed.

```{r}
# Reshape using pivot_longer
known_long <- known %>%
  pivot_longer(
    cols = c(skull_length, skull_height, skull_width),  # Columns to reshape
    names_to = "measurement",                          # New column for measurement type
    values_to = "value"                                # New column for measurement values
  )

#Boxplot comparison in a single plot
ggplot(known_long, aes(x = chromosomal_id, y = value, fill=chromosomal_id)) +
  geom_boxplot() +
  facet_wrap(~ measurement, scales = "free",
             labeller = labeller(measurement = c("skull_length" = "Vole Skull Length",
                                                 "skull_height" = "Vole Skull height",
                                                 "skull_width" = "Vole Skull width"))) +
  labs(title = "Boxplot of Vole Skull Measurements by Chromosomal Id",
       x = "Chromosomal Id",
       y = "Measurement (0.01mm)") +
  theme_minimal()
```

## Histogram

Observe the known dataset in a histogram, confirming outliers were removed.

```{r}
#Histogram comparison in a single plot
ggplot(known_long, aes(x = value, fill = chromosomal_id)) +
  geom_histogram(position = "identity", alpha = 0.6, bins = 30) +  
  facet_wrap(~ measurement, scales = "free", 
             labeller = labeller(measurement = c("skull_length" = "Vole Skull Length",
                                                 "skull_height" = "Vole Skull Height",
                                                 "skull_width" = "Vole Skull Width"))) +
  labs(title = "Histogram of Vole Skull Measurements by Chromosomal Id",
       x = "Measurement (0.01mm)",
       y = "Frequency") +
  theme_minimal() +
  scale_fill_manual(values = c("subterraneus" = "blue", "multiplex" = "red"))

```

**Comment on the boxplot:** The outliers that are identified in the boxplot, but not by the Interquartile Range Threshold of (1\*IQR) i.e., not labeled, are most likely due to the differences in calculation of outliers by the boxplot. The outliers here can be considered minor deviation within the Threshold and are considered non-material.

## Pairwise correlation plot

Observe the known dataset in a correlation plot.

```{r}

# Create a pair plot, coloring by Chromosomal type
ggpairs(
  known,
  columns = c("skull_length", "skull_height", "skull_width"),
  aes(color = chromosomal_id, alpha = 0.5)
) + 
  labs(title = "Pairwise comparision") +
  theme_minimal()

```

# Data Manipulation for unknown Dataset

Removal of NA values in the unknown dataset that do not have a column in any of the columns (i.e., blank rows).

```{r}
#remove NA rows with no value in all of the columns
unknown <- unknown %>% filter(rowSums(is.na(.)) < ncol(.))

library(psych)
# Check the cleaned and renamed data set
cat("\nCleaned Unknown Data:\n")
describe(unknown)

```

### Visualize unknown dataset

Observe the unknown dataset in a histogram, viewing the outliers.

```{r, warning=FALSE}
# Create box plots of the known data to review outliers
unknown_skull_height <- ggplot(unknown, aes(y = skull_height)) + 
  geom_boxplot(fill = "lightblue", color = "darkblue") + 
  ggtitle("Box Plot for Skull Height") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))

unknown_skull_length <- ggplot(unknown, aes(y = skull_length)) + 
  geom_boxplot(fill = "lightgreen", color = "darkgreen") + 
  ggtitle("Box Plot for Skull Length") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))

unknown_skull_width <- ggplot(unknown, aes(y = skull_width)) + 
  geom_boxplot(fill = "lightcoral", color = "darkred") + 
  ggtitle("Box Plot for Skull Width") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))

# Combine all plots into one layout
library(gridExtra)
grid.arrange(unknown_skull_height, unknown_skull_length, unknown_skull_width, ncol = 3)
```

## Replace large outliers (typos) by capping

Replace outliers with the specified values based on the specified bounds (capping outliers). This was identified based on manual inspection, since the data set was small. We didn't see any typos for skull_height

-   For skull_length: replace values below 1907 with 1900 and values above 2606 with 2600
-   For skull_width: replace values below 374 with 400 and values above 546 with 500

**Note:** We are only targeting extremely large outliers that were identified as typo, while maintaining the variability in the data.

```{r}

# Initialize vectors to store indices and original values of replaced outliers (Which indices are True?)
skull_length_replaced <- which(!is.na(unknown$skull_length) & (unknown$skull_length < 1907 | unknown$skull_length > 2606))
skull_width_replaced <- which(!is.na(unknown$skull_width) & (unknown$skull_width < 374 | unknown$skull_width > 546))

# Capture original values that are outliers
original_skull_length <- unknown$skull_length[skull_length_replaced]
original_skull_width <- unknown$skull_width[skull_width_replaced]

# Replace outliers as specified
unknown$skull_length <- ifelse(!is.na(unknown$skull_length) & unknown$skull_length < 1907, 1900,
                                ifelse(!is.na(unknown$skull_length) & unknown$skull_length > 2606, 2600, unknown$skull_length))

unknown$skull_width <- ifelse(!is.na(unknown$skull_width) & unknown$skull_width < 374, 400,
                               ifelse(!is.na(unknown$skull_width) & unknown$skull_width > 546, 500, unknown$skull_width))

# Combine indices and original values into a data frame for easy inspection
replaced_values <- list(
  skull_length = data.frame(Index = skull_length_replaced, Outlier_Value = original_skull_length, replaced_Value = unknown$skull_length[skull_length_replaced]),
  skull_width = data.frame(Index = skull_width_replaced, Outlier_Value = original_skull_width, replaced_Value = unknown$skull_width[skull_width_replaced])
)

# Print the replaced values
print(replaced_values)

#Remove Chromosomal_id from unknown 
unknown<- unknown %>%
  select(-(`chromosomal_id`))
```

# EDA for Unknown Dataset

Observe the unknown dataset in a boxplot and descriptive summary, confirming outliers were removed.

```{r, warning=FALSE}
# Descriptive Summary after removing outliers
cat("\nDescriptive Summary:\n")
describe(unknown)

# Create box plots of the unknown data to verify outliers corrected
unknown_skull_height <- ggplot(unknown, aes(y = skull_height)) + 
  geom_boxplot(fill = "lightblue", color = "darkblue") + 
  ggtitle("Box Plot for Skull Height") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))

unknown_skull_length <- ggplot(unknown, aes(y = skull_length)) + 
  geom_boxplot(fill = "lightgreen", color = "darkgreen") + 
  ggtitle("Box Plot for Skull Length") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))

unknown_skull_width <- ggplot(unknown, aes(y = skull_width)) + 
  geom_boxplot(fill = "lightcoral", color = "darkred") + 
  ggtitle("Box Plot for Skull Width") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))

# Combine all plots into one layout
grid.arrange(unknown_skull_height, unknown_skull_length, unknown_skull_width, ncol = 3)
```

# Cross Validation and Prediction

## Model with all three variables (Model_LWH)

```{r}
#for reproducibility
set.seed(1234)

# to store data
predictions <- numeric(nrow(known))
actual <- known$response

# Leave One Out Cross Validation (LOOCV) on the known dataset
for (i in 1:nrow(known)){
  
  train_data <- known[-i, ]
  test_data <- known[i, ]
  
  model_LWH <- glm(response ~ skull_length + skull_height + skull_width, data = train_data, family = 'binomial')
  
  predictions[i] <- predict(model_LWH, newdata = test_data, type = "response")
  
}

# check the model
summary(model_LWH)

# Odds Ratios
exp(coef(model_LWH))

# Count of each species
table(known$response)

# Proportion of each species
prop.table(table(known$response))

#MSE
mse <- mean((predictions - actual)^2)  # Mean Squared Error
print(paste("LOOCV Mean Squared Error:", mse))

#Confusion matrix
predicted_class <- ifelse(predictions > 0.5, 1, 0)  # Threshold at 0.5
accuracy <- mean(predicted_class == actual)
print(paste("Accuracy:", accuracy))

print(paste("Confusion matrix for Model_LWH"))
table(Predicted = predicted_class, Actual = actual)

#########################################################

#Create new dataframe to store predictions from all model
Model_prediction <- unknown %>% 
  as.data.frame()

#Prediction on unknown
probabilities <- predict(model_LWH, newdata = unknown, type = "response")
#convert predicted probabilities to prediction class with threshold at 0.5
unknown$Predicted_response <- ifelse(probabilities > 0.5, 1, 0)
# identify chromosomal id where 1 represents subterraneus and 0 represents multiplex
Model_prediction$Predict_LWH<- ifelse(unknown$Predicted_response == 1, "subterraneus", "multiplex")

#Count the number of prediction
print(paste("Prediction count"))
table(Model_prediction$Predict_LWH)

```

## Model with two variables (Model_LH)

```{r}
#for reproducibility
set.seed(1234)

# to store data
predictions <- numeric(nrow(known))
actual <- known$response

# Leave One Out Cross Validation (LOOCV) on the known dataset
for (i in 1:nrow(known)){
  
  train_data <- known[-i, ]
  test_data <- known[i, ]
  
  model_LH <- glm(response ~ skull_length + skull_height, data = train_data, family = 'binomial')
  
  predictions[i] <- predict(model_LH, newdata = test_data, type = "response")
  
}

# check the model
summary (model_LH)

#MSE
mse <- mean((predictions - actual)^2)  # Mean Squared Error
print(paste("LOOCV Mean Squared Error:", mse))

#Confusion matrix
predicted_class <- ifelse(predictions > 0.5, 1, 0)  # Threshold at 0.5
accuracy <- mean(predicted_class == actual)
print(paste("Accuracy:", accuracy))

print(paste("Confusion matrix for Model_LH"))
table(Predicted = predicted_class, Actual = actual)

#########################################################

#Prediction on unknown
probabilities <- predict(model_LH, newdata = unknown, type = "response")
#convert predicted probabilities to prediction class with threshold at 0.5
unknown$Predicted_response <- ifelse(probabilities > 0.5, 1, 0)
# identify chromosomal id where 1 represents subterraneus and 0 represents multiplex
Model_prediction$Predict_LH<- ifelse(unknown$Predicted_response == 1, "subterraneus", "multiplex")

#Count the number of prediction
print(paste("Prediction count"))
table(Model_prediction$Predict_LH)

```

## Model with two variables (Model_HW)

```{r}
#for reproducibility
set.seed(1234)

# to store data
predictions <- numeric(nrow(known))
actual <- known$response

# Leave One Out Cross Validation (LOOCV) on the known dataset
for (i in 1:nrow(known)){
  
  train_data <- known[-i, ]
  test_data <- known[i, ]
  
  model_HW <- glm(response ~ skull_height + skull_width, data = train_data, family = 'binomial')
  
  predictions[i] <- predict(model_HW, newdata = test_data, type = "response")
  
}

# check the model
summary (model_HW)

#MSE
mse <- mean((predictions - actual)^2)  # Mean Squared Error
print(paste("LOOCV Mean Squared Error:", mse))

#Confusion matrix
predicted_class <- ifelse(predictions > 0.5, 1, 0)  # Threshold at 0.5
accuracy <- mean(predicted_class == actual)
print(paste("Accuracy:", accuracy))

print(paste("Confusion matrix for Model_HW"))
table(Predicted = predicted_class, Actual = actual)

#########################################################

#Prediction on unknown
probabilities <- predict(model_HW, newdata = unknown, type = "response")
#convert predicted probabilities to prediction class with threshold at 0.5
unknown$Predicted_response <- ifelse(probabilities > 0.5, 1, 0)
# identify chromosomal id where 1 represents subterraneus and 0 represents multiplex
Model_prediction$Predict_HW<- ifelse(unknown$Predicted_response == 1, "subterraneus", "multiplex")

#Count the number of prediction
print(paste("Prediction count"))
table(Model_prediction$Predict_HW)

```

## Model with two variables (Model_LW)

```{r}
#for reproducibility
set.seed(1234)

# to store data
predictions <- numeric(nrow(known))
actual <- known$response

# Leave One Out Cross Validation (LOOCV) on the known dataset
for (i in 1:nrow(known)){
  
  train_data <- known[-i, ]
  test_data <- known[i, ]
  
  model_LW <- glm(response ~ skull_length + skull_width, data = train_data, family = 'binomial')
  
  predictions[i] <- predict(model_LW, newdata = test_data, type = "response")
  
}

# check the model
summary (model_LW)

#MSE
mse <- mean((predictions - actual)^2)  # Mean Squared Error
print(paste("LOOCV Mean Squared Error:", mse))

#Confusion matrix
predicted_class <- ifelse(predictions > 0.5, 1, 0)  # Threshold at 0.5
accuracy <- mean(predicted_class == actual)
print(paste("Accuracy:", accuracy))

print(paste("Confusion matrix for Model_LW"))
table(Predicted = predicted_class, Actual = actual)

#########################################################

#Prediction on unknown
probabilities <- predict(model_LW, newdata = unknown, type = "response")
#convert predicted probabilities to prediction class with threshold at 0.5
unknown$Predicted_response <- ifelse(probabilities > 0.5, 1, 0)
# identify chromosomal id where 1 represents subterraneus and 0 represents multiplex
Model_prediction$Predict_LW<- ifelse(unknown$Predicted_response == 1, "subterraneus", "multiplex")

#Count the number of prediction
print(paste("Prediction count"))
table(Model_prediction$Predict_LW)

```

## Model with one variable (Model_L)

```{r}
#for reproducibility
set.seed(1234)

# to store data
predictions <- numeric(nrow(known))
actual <- known$response

# Leave One Out Cross Validation (LOOCV) on the known dataset
for (i in 1:nrow(known)){
  
  train_data <- known[-i, ]
  test_data <- known[i, ]
  
  model_L <- glm(response ~ skull_length, data = train_data, family = 'binomial')
  
  predictions[i] <- predict(model_L, newdata = test_data, type = "response")
  
}

# check the model
summary (model_L)

#MSE
mse <- mean((predictions - actual)^2)  # Mean Squared Error
print(paste("LOOCV Mean Squared Error:", mse))

#Confusion matrix
predicted_class <- ifelse(predictions > 0.5, 1, 0)  # Threshold at 0.5
accuracy <- mean(predicted_class == actual)
print(paste("Accuracy:", accuracy))

print(paste("Confusion matrix for Model_L"))
table(Predicted = predicted_class, Actual = actual)

#########################################################

#Prediction on unknown
probabilities <- predict(model_L, newdata = unknown, type = "response")
#convert predicted probabilities to prediction class with threshold at 0.5
unknown$Predicted_response <- ifelse(probabilities > 0.5, 1, 0)
# identify chromosomal id where 1 represents subterraneus and 0 represents multiplex
Model_prediction$Predict_L<- ifelse(unknown$Predicted_response == 1, "subterraneus", "multiplex")

#Count the number of prediction
print(paste("Prediction count"))
table(Model_prediction$Predict_L)


```

## Model with one variable (Model_H)

```{r}
#for reproducibility
set.seed(1234)

# to store data
predictions <- numeric(nrow(known))
actual <- known$response

# Leave One Out Cross Validation (LOOCV) on the known dataset
for (i in 1:nrow(known)){
  
  train_data <- known[-i, ]
  test_data <- known[i, ]
  
  model_H <- glm(response ~ skull_height, data = train_data, family = 'binomial')
  
  predictions[i] <- predict(model_H, newdata = test_data, type = "response")
  
}

# check the model
summary (model_H)

#MSE
mse <- mean((predictions - actual)^2)  # Mean Squared Error
print(paste("LOOCV Mean Squared Error:", mse))

#Confusion matrix
predicted_class <- ifelse(predictions > 0.5, 1, 0)  # Threshold at 0.5
accuracy <- mean(predicted_class == actual)
print(paste("Accuracy:", accuracy))

print(paste("Confusion matrix for Model_H"))
table(Predicted = predicted_class, Actual = actual)

#########################################################

#Prediction on unknown
probabilities <- predict(model_H, newdata = unknown, type = "response")
#convert predicted probabilities to prediction class with threshold at 0.5
unknown$Predicted_response <- ifelse(probabilities > 0.5, 1, 0)
# identify chromosomal id where 1 represents subterraneus and 0 represents multiplex
Model_prediction$Predict_H<- ifelse(unknown$Predicted_response == 1, "subterraneus", "multiplex")

#Count the number of prediction
print(paste("Prediction count"))
table(Model_prediction$Predict_H)

```

## Model with one variable (Model_W)

```{r}
#for reproducibility
set.seed(1234)

# to store data
predictions <- numeric(nrow(known))
actual <- known$response

# Leave One Out Cross Validation (LOOCV) on the known dataset
for (i in 1:nrow(known)){
  
  train_data <- known[-i, ]
  test_data <- known[i, ]
  
  model_W <- glm(response ~ skull_width, data = train_data, family = 'binomial')
  
  predictions[i] <- predict(model_W, newdata = test_data, type = "response")
  
}

# check the model
summary (model_W)

#MSE
mse <- mean((predictions - actual)^2)  # Mean Squared Error
print(paste("LOOCV Mean Squared Error:", mse))

#Confusion matrix
predicted_class <- ifelse(predictions > 0.5, 1, 0)  # Threshold at 0.5
accuracy <- mean(predicted_class == actual)
print(paste("Accuracy:", accuracy))

print(paste("Confusion matrix for Model_W"))
table(Predicted = predicted_class, Actual = actual)

#########################################################

#Prediction on unknown
probabilities <- predict(model_W, newdata = unknown, type = "response")
#convert predicted probabilities to prediction class with threshold at 0.5
unknown$Predicted_response <- ifelse(probabilities > 0.5, 1, 0)
# identify chromosomal id where 1 represents subterraneus and 0 represents multiplex
Model_prediction$Predict_W<- ifelse(unknown$Predicted_response == 1, "subterraneus", "multiplex")

#Count the number of prediction
print(paste("Prediction count"))
table(Model_prediction$Predict_W)

```

### Prediction count

```{r}
# Replace missing value with string NA

Model_prediction[] <- lapply(Model_prediction, function(x) {
  ifelse(is.na(x), "NA", x)
})

# Count occurrences for each column
counts <- sapply(Model_prediction[, grep("^Predict_", names(Model_prediction))], 
                 function(x) table(factor(x, levels = c("subterraneus", "multiplex", "NA"))))
print(paste("Prediction counts by each model:"))
counts

```

# Final Prediction

We made the Final Prediction based on the highest frequency of predicted chromosomal type for each index.

```{r}
library(dplyr)

library(dplyr)

Model_prediction <- Model_prediction %>%
  # Count occurrences of multiplex and subterraneus for each row
  rowwise() %>%
  mutate(
    multiplex_count = sum(c_across(starts_with("Predict_")) == "multiplex", na.rm = TRUE),
    subterraneus_count = sum(c_across(starts_with("Predict_")) == "subterraneus", na.rm = TRUE),
    # Create Final_Prediction based solely on frequency counts
    Final_Prediction = if_else(multiplex_count > subterraneus_count, "multiplex", "subterraneus")
  ) %>%
  ungroup()

#Count the number of prediction
print(paste("Prediction count"))
table(Model_prediction$Final_Prediction)

# View the final updated table
print(Model_prediction)

#export the excel file
library(writexl)
write_xlsx(Model_prediction, path = "VoleSkull_predicted.xlsx")
```

## Visualize Final Prediction

```{r}
# Bar plot of Final_Prediction counts with labels
ggplot(Model_prediction, aes(x = Final_Prediction, fill = Final_Prediction)) +
  geom_bar() +
  geom_text(
    stat = 'count', 
    aes(label = ..count..), 
    vjust = -0.5, 
    size = 4
  ) +
  labs(
    title = "Count of Final Predicted Chromosomal ID",
    x = "Chromosomal ID",
    y = "Count"
  ) +
  theme_minimal()

```
